---
title: "Building Tidy Data"
author: "Ian Dinwoodie"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(echo=TRUE)
```

# Assemble the Study Data Frame

## Dog Data Processing

### Load Raw Data

Load the raw dog data and verify its dimensions and structure. We expect to have
1704 responses across 29 fields.

```{r}
df_dog <- read.csv('../data/raw/dogs.csv', header=TRUE, skipNul=TRUE)
str(df_dog, list.len=5)
stopifnot(identical(dim(df_dog)+0, c(1704, 29))) # Verify expected dim.
```

### Apply Readable Column Names

A list of improved column names focused on readability has been provided in the
reference directory. Let's load and apply these updated names to our data frame.

```{r}
colnames(df_dog) <- readr::read_lines(
  '../references/dog-readable-column-names.txt')
str(df_dog)
```

### Drop Unnecessary Columns

We can drop the `internal_logic_*` columns since their only purpose was for
Typeform UI enhancement at the time of the survey. This operation will drop 5
fields from the data frame.

```{r}
df_dog <- df_dog %>%
  select(-starts_with("internal_logic_"))
stopifnot(identical(dim(df_dog)+0, c(1704, 24))) # Verify 5 columns dropped.
```

### Adjust Data Types

```{r}
# Convert logical fields.
df_dog <- df_dog %>%
  mutate_at(c("is_living_with_dog", "is_open_to_another_dog"), as.logical)

# Convert factor fields.
factors <- c(
  'time_thinking',
  'primary_motivation',
  'acquisition_source',
  'important_char_1',
  'important_char_2',
  'important_char_3',
  'important_char_4',
  'important_char_5',
  'important_char_6',
  'met_expectations',
  'time_together_len',
  'curr_dog_location',
  'new_important_char_1',
  'new_important_char_2',
  'new_important_char_3',
  'new_important_char_4',
  'new_important_char_5',
  'new_important_char_6',
  'new_acquisition_source'
)
df_dog <- df_dog %>%
  mutate_at(factors, as.factor)
rm(factors)

# Adjust factor levels.
primary_motivations <- c(
  "Companionship and affection",
  "Social interaction",
  "Exercise, adventure partner",
  "Protection",
  "Working or sporting",
  "Someone else in the home wanted a dog (e.g. kids, spouse)",
  "Companion for another dog or pet"
)
df_dog$primary_motivation <- as.factor(ifelse(
  df_dog$primary_motivation %in% primary_motivations,
  as.character(df_dog$primary_motivation),
  "other"
))
rm(primary_motivations)
levels(df_dog$primary_motivation) <- c(
  "companionship",
  "social interaction",
  "exercise",
  "protection",
  "working",
  "someone else",
  "trainability",
  "friend for pet"
)
simple_levels <- c(
  "age",
  "appearance",
  "breed",
  "compatability",
  "personality",
  "size",
  "trainability"
)
levels(df_dog$important_char_1) <- simple_levels
levels(df_dog$important_char_2) <- simple_levels
levels(df_dog$important_char_3) <- simple_levels
levels(df_dog$important_char_4) <- simple_levels
levels(df_dog$important_char_5) <- simple_levels
levels(df_dog$important_char_6) <- simple_levels
simple_levels <- c("", simple_levels)
levels(df_dog$new_important_char_1) <- simple_levels
levels(df_dog$new_important_char_2) <- simple_levels
levels(df_dog$new_important_char_3) <- simple_levels
levels(df_dog$new_important_char_4) <- simple_levels
levels(df_dog$new_important_char_5) <- simple_levels
levels(df_dog$new_important_char_6) <- simple_levels
rm(simple_levels)

# data.frame(levels = unique(df_dog$time_thinking), value = as.numeric(unique(df_dog$time_thinking)))
# df_dog$time_thinking <- factor(
#   df_dog$time_thinking,
#   levels=c(2,3,4,1),
#   labels= c("1wk-6mos", "6mos-6yrs", ">6yrs", "<1wk")
# )
# data.frame(levels = unique(df_dog$time_thinking), value = as.numeric(unique(df_dog$time_thinking)))

#simple_levels <- c(
#  "<1wk",
#  "1wk-6mos",
#  "6mos-6yrs",
#  ">6yrs"
#)
#df_dog$time_thinking <- factor(df_dog$time_thinking, levels=simple_levels,
#                               ordered=TRUE)

# levels(df_dog$time_together_len) <- c(
#   "",
#   "1wk-6mos",
#   "6mos-6yrs",
#   ">6yrs",
#   "<1wk"
# )
levels(df_dog$acquisition_source) <- c(
  "import",
  "breeder",
  "family-friend",
  "found",
  "pet shop",
  "online",
  "shelter"
)
levels(df_dog$new_acquisition_source) <- c(
  "",
  "import",
  "breeder",
  "family-friend",
  "found",
  "pet shop",
  "online",
  "shelter"
)
levels(df_dog$met_expectations) <- c(
  "no",
  "partially-time",
  "partially-acceptance",
  "yes"
)
dog_locations <- c(
  "Rehomed",
  "Surrendered to shelter/rescue",
  "Euthanized",
  "Ran away/got lost",
  "Passed",
  ""
)
df_dog$curr_dog_location <- as.factor(ifelse(
  df_dog$curr_dog_location %in% dog_locations,
  as.character(df_dog$curr_dog_location),
  "other"
))
rm(dog_locations)
levels(df_dog$curr_dog_location) <- c(
  "",
  "euthanized",
  "other",
  "passed",
  "lost",
  "rehomed"
)

# Convert date fields.
df_dog$datetime <- as.POSIXct(strptime(df_dog$datetime,"%m/%d/%Y %H:%M:%S"))

# Normalize char fields.
df_dog$name <- trimws(tolower(df_dog$name))

# Examine the updated data frame.
str(df_dog)
```

### Handle Repeat Responses

We can examine the (`owner_id`, `name`) pairs, which we will refer to as the
unique response identifier, in the data frame to determine if there are repeat
responses for any of the dogs.

```{r}
df_dog %>%
  group_by(owner_id, name) %>%
  count() %>%
  filter(n>1) %>%
  arrange(desc(n))
```

We see that there are 109 non-unique response identifiers and that the maximum
occurrence of an identifier is 2. Therefore, elimination of these duplicates
should yield a data frame with $1704-109=1595$ responses.

```{r}
df_dog <- df_dog %>%
  group_by(owner_id, name) %>%
  slice_max(datetime) %>%
  ungroup()
dim(df_dog)
stopifnot(identical(dim(df_dog)+0, c(1595, 24))) # Verify expected dim.
```

We see that the number of responses in the data frame now matches the number of
unique response identifiers for this data frame.

### Summary

Let's look at a summary of the dog data before we move on.

```{r}
summary(df_dog)
```

## Owner Data Processing

### Load Raw Data

Load the raw owner data and verify its dimensions and structure. We expect to
have 1225 responses across 6 fields.

```{r}
df_owner <- read.csv('../data/raw/owners.csv', header=TRUE, skipNul=TRUE)
str(df_owner)
stopifnot(identical(dim(df_owner)+0, c(1225, 6))) # Verify expected dim.
```

### Apply Readable Column Names

A list of improved column names focused on readability has been provided in the
reference directory. Let's load and apply these updated names to our data frame.

```{r}
colnames(df_owner) <- readr::read_lines(
  '../references/owner-readable-column-names.txt')
str(df_owner)
```

### Drop Unnecessary Columns

We don't need to retain the `internal_logic_1` column since its only purpose was
Typeform internal use nor do we need the `zip_code`; it is an unreliable
measure since not all respondents are located in the US.

```{r}
df_owner <- subset(df_owner, select=-c(
  internal_logic_1,
  zip_code
))
stopifnot(identical(dim(df_owner)+0, c(1225, 4))) # Verify 2 columns dropped.
```

### Adjust Data Types

```{r}
# Convert numeric and logical fields.
df_owner <- df_owner %>%
  mutate_at("age", as.numeric) %>%
  mutate_at("sex", as.factor)

# Convert date fields.
df_owner$datetime <- as.POSIXct(strptime(df_owner$datetime,"%m/%d/%Y %H:%M:%S"))

# Examine the updated data frame.
str(df_owner)
```

### Handle Repeat Responses

For this data frame the `owner_id` is the unique response identifier. We expect
one response for each `owner_id`;  let's verify this expectation.

```{r}
df_owner %>%
  group_by(owner_id) %>%
  count() %>%
  filter(n>1) %>%
  arrange(desc(n))
```

We see that there are 154 non-unique response identifiers and that the maximum
occurrence of an identifier is 4. Therefore, elimination of these duplicates
should yield a data frame with $1225-154-7-1=1063$ responses.

```{r}
df_owner <- df_owner %>%
  group_by(owner_id) %>%
  slice_max(datetime) %>%
  ungroup()
dim(df_owner)
stopifnot(identical(dim(df_owner)+0, c(1063, 4))) # Verify expected dim.
```

We see that the number of responses in the data frame now matches the number of
unique response identifiers for this data frame.

### Summary

Let's look at a summary of the owner data before we move on.

```{r}
summary(df_owner)
```

## Build Unified Data Frame

We now use both the dog and owner data frames to construct a unified data frame.
This is the data frame that will be used for all analyses in the study.

### Verifying Agreement of Owner Unique Identifiers

The set of owner ID's in the dog data frame (set A) must be a subset of the set
of owner ID's in the owner data frame (set B). We can verify this by evaluating
the equality: $A \cap B = A$.

```{r}
set_a <- df_dog %>%
  select(owner_id) %>%
  unique()
set_b <- df_owner %>%
  select(owner_id) %>%
  unique()
result <- setequal(intersect(set_a, set_b), set_a)
print(paste("A ∩ B = A:", result))
stopifnot(result)
rm(set_a, set_b, result)
```

### Merge Data Frames

We merge the dog and owner data frames about the owner ID's and expect the
resulting data frame to have 1595 responses (the same number of responses as the
dog data frame) across 27 fields ($24+4-1=27$).

```{r}
df <- merge(df_dog, df_owner, by="owner_id")
str(df)
stopifnot(identical(dim(df)+0, c(1595, 27))) # Verify expected dim.
rm(df_dog, df_owner)
```

### Drop Unnecessary Columns

```{r}
df <- df %>%
  select(-starts_with("datetime"))
stopifnot(identical(dim(df)+0, c(1595, 25))) # Verify 2 columns dropped.
```

# Derived Columns

## Generate Rank Columns

```{r}
df$age_rank <- 7
df$appearance_rank <- 7
df$breed_rank <- 7
df$compatability_rank <- 7
df$personality_rank <- 7
df$size_rank <- 7
df$trainability_rank <- 7

char_cols <- c(
  'important_char_1',
  'important_char_2',
  'important_char_3',
  'important_char_4',
  'important_char_5',
  'important_char_6'
)

for (c in char_cols) {
  rank <- which(char_cols == c)
  for (r in 1:nrow(df)) {
    rank_col <- paste0(as.character(df[r, c]), "_rank")
    df[r, rank_col] <- rank
  }
}
rm(char_cols, c, r, rank, rank_col)

str(df)
stopifnot(identical(dim(df)+0, c(1595, 32))) # Verify 7 columns added.
```

## Generate Satisfaction Columns

```{r}
#df$is_satisfied <- ifelse(
#  df$met_expectations == "No, the dog was not a good match",
#  FALSE,
#  TRUE
#)
df$is_satisfied <- ifelse(
  df$met_expectations == "Yes",
  TRUE,
  FALSE
)

df$satisfaction_score <- as.numeric(df$met_expectations)

stopifnot(identical(dim(df)+0, c(1595, 34))) # Verify 2 columns added.
```

# Drop Low Ages

```{r}
#df <- df %>%
#  filter(age >= 18)
#stopifnot(identical(dim(df)+0, c(1543, 34))) # Verify rows dropped.

```

# Final Summary

Take a last look at the data before saving it to disk.

```{r}
dim(df)
summary(df)
```

# Saving the Tidy Data

Save the data to a file in RDS format so that the data types are saved and the
output file is compressed.

```{r}
saveRDS(df, '../data/processed/tidy.Rds')
```
