---
title: "Statistical Inference"
author: "Ian Dinwoodie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(MASS)
library(caret)
library(car)
library(performance)

knitr::opts_chunk$set(echo=TRUE)
set.seed(1)
```

# Preparing the Data

## Loading the Tidy Data

Load the tidy data from disk.

```{r}
df <- readRDS('../data/processed/tidy.Rds')
stopifnot(identical(dim(df)+0, c(1537, 29)))
str(df, list.len=5)
```

## Drop/Rename/Reorder Columns

Drop or adjust (rename/reorder) columns.

```{r}
old_names <- c(
  colnames(df %>% dplyr::select(contains("rank") & !contains("revised"))),
  "time_thinking", "primary_motivation", "acquisition_source",
  "is_living_with_dog")
new_names <- c(
  "ch.age", "ch.appearance", "ch.breed", "ch.compatability", "ch.personality",
  "ch.size", "ch.trainability", "think_len", "motive", "source",
  "is_with_dog")

df <- df %>%
  dplyr::rename_with(~ new_names, all_of(old_names)) %>%
  dplyr::select(
    -owner_id,
    -dog_name,
    -met_expectations,
    -time_together_len,
    -curr_dog_location,
    -is_consider_another_dog,
    -is_with_dog,
    -contains("revised"),
    -owner_gender
  ) %>%
  dplyr::relocate(is_owner_male, .before = everything()) %>%
  dplyr::relocate(owner_age , .after=is_owner_male) %>%
  dplyr::relocate(is_satisfied, .after = last_col()) %>%
  tidyr::drop_na()

rm(old_names, new_names)
dim(df)
summary(df)
stopifnot(identical(dim(df)+0, c(1537, 13)))
```

# Preprocessing

## Rule of Ten

We start by identifying any "rule of ten" violations:

```{r}
table(df$is_satisfied, df$is_owner_male)
```

```{r}
table(df$is_satisfied, df$think_len)
```

```{r}
table(df$is_satisfied, df$motive)
```

We see violations for motives `protection` and `social`.

```{r}
table(df$is_satisfied, df$source)
```

We see violations for motives `protection` and `social` and for sources
`foreign` and `found`. These columns will need to be dropped from the
preprocessed data.

## Establish Reference Levels

```{r}
df <- df %>%
  dplyr::mutate(motive = forcats::fct_relevel(motive, "companionship")) %>%
  dplyr::mutate(source = forcats::fct_relevel(source, "pet_shop"))
summary(df)
```

## Create Preprocessed Data

Create dummy variables.

```{r}
# Reference: https://bookdown.org/rehk/stm1001_dsm_introduction_to_machine_learning_in_r/machine-learning-in-r-using-the-caret-package.html
dummies <- caret::dummyVars(is_satisfied ~ ., data = df, fullRank = T)
df_pp <- cbind(as_tibble(predict(dummies, newdata = df)),
               is_satisfied=df$is_satisfied)
df_pp <- df_pp %>%
  dplyr::rename(is_owner_male=is_owner_maleTRUE) %>%
  # Drop rule of ten violations.
  dplyr::select(
    -c(motive.protection, motive.social, source.found, source.foreign))
head(df_pp)
```

Check for near-zero variance variables.

```{r}
(nzv_metrics <- caret::nearZeroVar(df_pp, saveMetrics=TRUE))
nzv_cnt <- sum(nzv_metrics$nzv)
print(paste("Warning:", nzv_cnt, "near-zero variance vars. found"))
if (nzv_cnt) {
  print(colnames(df_pp[, nzv_metrics$nzv]))
}
```

Check for linearly correlated predictors.

```{r}
combo_info <- caret::findLinearCombos(
  df_pp %>% dplyr::select(-contains("ch."), -is_satisfied))
combo_cnt <- length(combo_info$remove)
print(paste("Linearly corr. combos vars to examine:", combo_cnt))
print(paste("Warning:", combo_cnt, "linearly corr. combos vars. found"))
if (combo_cnt) {
  print(colnames(df_pp)[combo_info$remove])
}
```

## Modified

Instead of discarding the near-zero variance variables, we collapse fields or
portions of fields and check to see if the diagnostics are improved.

```{r}
# dummies <- caret::dummyVars(is_satisfied ~ ., data = df2, fullRank=T)
# df_pp <- cbind(as_tibble(predict(dummies, newdata = df2)),
#                is_satisfied=df2$is_satisfied)
# df_pp <- df_pp %>%
#   dplyr::rename(is_owner_male=is_owner_maleTRUE)
# head(df_pp)
# 
# (nzv_metrics <- caret::nearZeroVar(df_pp, saveMetrics=TRUE))
# nzv_cnt <- sum(nzv_metrics$nzv)
# print(paste("Warning:", nzv_cnt, "near-zero variance vars. found"))
# if (nzv_cnt) {
#   print(colnames(df_pp[, nzv_metrics$nzv]))
# }
```

Check for linearly correlated predictors.

```{r}
# combo_info <- caret::findLinearCombos(
#   df_pp %>% dplyr::select(-contains("ch."), -is_satisfied))
# combo_cnt <- length(combo_info$remove)
# print(paste("Linearly corr. combos vars to examine:", combo_cnt))
# print(paste("Warning:", combo_cnt, "linearly corr. combos vars. found"))
# if (combo_cnt) {
#   print(colnames(df_pp)[combo_info$remove])
# }
```

```{r}
str(df_pp)
```

# Models

## Null Model

```{r}
glm_fit_null <- glm(is_satisfied~1, df_pp, family="binomial")
summary(glm_fit_null)

(r2_null <- performance::r2_nagelkerke(glm_fit_null))
```

## Background Variables

```{r}
df_back <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age, contains("think_len."))
summary(df_back)
```

```{r}
set.seed(1)
glm_fit_back <- glm(is_satisfied ~ ., df_back, family="binomial")
summary(glm_fit_back)

get_performance_metrics <- function(fit, r2_prev) {
  r2_curr = performance::r2_nagelkerke(fit)
  print(r2_curr)
  print(paste("Delta R^2:", r2_curr-r2_prev))
  print(performance::performance_hosmer(fit))
  return(r2_curr)
}

r2_back <- get_performance_metrics(glm_fit_back, r2_null)
```

```{r}
(vif(glm_fit_back))
```

```{r}
bs_table <- function(fit, data) {
  df_results <- broom::tidy(fit, conf.int=TRUE, exponentiate=TRUE)
  
  glm_est <- function(split, ...) {
    glm(fit$formula, data = rsample::analysis(split), family="binomial") %>%
      tidy()
  }
  
  set.seed(1)
  conf_ints <- rsample::bootstraps(data, 10000, apparent = TRUE) %>%
    dplyr::mutate(results = map(splits, glm_est)) %>%
    rsample::int_bca(results, .fn = glm_est) %>%
    dplyr::mutate_at(c(".estimate", ".lower", ".upper"), ~ exp(.)) %>%
    dplyr::select(term, .lower, .upper) %>%
    dplyr::rename(bs.low=.lower, bs.high=.upper)
  
  df_results <- merge(df_results, conf_ints, by="term")
  df_results$p.value <- p.adjust(df_results$p.value, method='fdr')
  df_results$sig <- ''
  df_results[df_results$p.value <= .05, 'sig'] <- '*'
  df_results[df_results$p.value <= .01, 'sig'] <- '**'
  df_results[df_results$p.value <= .001, 'sig'] <- '***'
  for (i in 1:nrow(df_results)) {
    if (is.na(df_results[i, 'bs.low']) | is.na(df_results[i, 'bs.high'])) next
    if ((df_results[i, 'bs.low'] <= 1) & (df_results[i, 'bs.high'] >= 1)) {
      df_results[i, 'sig'] <- ''
    }
  }
  df_results <- df_results %>%
    dplyr::mutate(dplyr::across(where(is.numeric), round, 3))
  return(knitr::kable(df_results))
}

bs_table(glm_fit_back, df_back)
```

```{r}
anova(glm_fit_null, glm_fit_back, test="LRT")
```

## Source

```{r}
df_source <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age, contains("think_len."),
                contains("source."))
summary(df_source)
```

```{r}
set.seed(1)
glm_fit_source <- glm(is_satisfied ~ ., df_source, family="binomial")
summary(glm_fit_source)

r2_source <- get_performance_metrics(glm_fit_source, r2_back)
```

```{r}
(vif(glm_fit_source))
```

```{r}
bs_table(glm_fit_source, df_source)
```

```{r}
anova(glm_fit_back, glm_fit_source, test="LRT")
```

## Motive

```{r}
df_motive <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."))
summary(df_motive)
```

```{r}
set.seed(1)
glm_fit_motive <- glm(is_satisfied~., df_motive, family="binomial")
summary(glm_fit_motive)

r2_motive <- get_performance_metrics(glm_fit_motive, r2_source)
```

```{r}
(vif(glm_fit_motive))
```

```{r}
bs_table(glm_fit_motive, df_motive)
```

```{r}
anova(glm_fit_source, glm_fit_motive, test="LRT")
```

## Characteristics

### Age

```{r}
char <- "ch.age"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Appearance

```{r}
char <- "ch.appearance"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Breed

```{r}
char <- "ch.breed"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Compatability

```{r}
char <- "ch.compatability"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Personality

```{r}
char <- "ch.personality"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Size

```{r}
char <- "ch.size"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

### Trainability

```{r}
char <- "ch.trainability"
df_char <- df_pp %>%
  dplyr::select(is_satisfied, is_owner_male, owner_age,
                contains("think_len."), contains("source."), contains("motive."),
                !!!char)
summary(df_char)
```

```{r}
set.seed(1)
glm_fit_char <- glm(is_satisfied~., df_char, family="binomial")
summary(glm_fit_char)

r2_char <- get_performance_metrics(glm_fit_char, r2_motive)
```

```{r}
(vif(glm_fit_char))
```

```{r}
bs_table(glm_fit_char, df_char)
```

```{r}
anova(glm_fit_motive, glm_fit_char, test="Chisq")
```

# Save Session Info

```{r}
sessionInfo()
```

# Cleanup

```{r}
rm(list = ls())
```
