---
title: "Exploratory Data Analysis (EDA)"
author: "Ian Dinwoodie"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(janitor)
library(psych)
library(caret)
library(vcd)
library(ggridges)
library(plotrix)

library(devtools)
devtools::install_github("vqv/ggbiplot")
library(ggbiplot)
devtools::install_github("kevinblighe/PCAtools")
library(PCAtools)
devtools::install_github("haleyjeppson/ggmosaic")
library(ggmosaic)
devtools::install_github("psyteachr/introdataviz")
library(introdataviz)

knitr::opts_chunk$set(echo=TRUE)
set.seed(1)
```

# Loading the Tidy Data

Load the tidy data from disk if it is available. If it is not available, run
`notebook-0-tidy.Rmd` to create it.

```{r}
df <- readRDS('../data/processed/tidy.Rds')
str(df)
stopifnot(identical(dim(df)+0, c(1537, 29)))
```

A summary of the loaded data is provided below.

```{r}
summary(df)
```

# Descriptive Stats

## Owner Demographics

The number of participating owners is equal to the number of unique owner
identifiers.

```{r}
length(unique(df$owner_id))
```

Generate a breakdown of owner gender.

```{r}
df %>%
  dplyr::distinct(owner_id, .keep_all=TRUE) %>%
  dplyr::count(owner_gender) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Generate summary statistics for owner ages.

```{r}
summary(
  df %>%
    dplyr::distinct(owner_id, .keep_all=TRUE) %>%
    dplyr::select(owner_age)
)
```

Examine the distribution of owner ages.

```{r}
df %>%
  ggplot(aes(x=owner_age)) +
    geom_histogram(alpha=0.5, position="identity", aes(y = ..density..),
                   color="black", bins=30) +
    geom_density() +
    geom_vline(xintercept=mean(df$owner_age), color="black", linetype="dashed",
               size=1)
```

Examine the distribution of ages split by owner gender.

```{r}
df %>%
  dplyr::select(is_owner_male, owner_age) %>%
  dplyr::transmute(male_age=ifelse(is_owner_male==TRUE, owner_age, NA),
                   female_age=ifelse(is_owner_male==FALSE, owner_age, NA)) %>%
  summary()
```

```{r}
means <- df %>%
  dplyr::group_by(owner_gender) %>%
  dplyr::summarise(means = mean(owner_age))
df %>%
  ggplot(aes(owner_age, fill = owner_gender)) +
    geom_histogram(alpha=0.7, position="identity", aes(y = ..density..),
                   color="black", bins=30) +
    geom_density(alpha=0.7) +
    geom_vline(xintercept=means$means, color=c("#F8766D", "#00BFC4"),
               linetype="dashed", size=0.75)
```

## Dog Demographics

The number of dogs corresponds to the number of rows in the data frame.

```{r}
dim(df)[1]
```

Generate summary statistics for the number of dogs per household.

```{r}
summary(as.data.frame(table(df$owner_id))$Freq)
```

Determine the number of single dog households.

```{r}
df %>%
  dplyr::count(owner_id) %>%
  dplyr::mutate(freq = n / sum(n)) %>%
  dplyr::filter(n == 1) %>%
  dplyr::mutate(freq = round(sum(freq) * 100, 2)) %>%
  dplyr::count(freq)
```
Generate breakdown of living situation.

```{r}
df %>%
  dplyr::count(is_living_with_dog) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

### Dogs Not-Living-With-Owner Demographics

```{r}
df %>%
  dplyr::filter(is_living_with_dog == FALSE) %>%
  dplyr::count(time_together_len) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

```{r}
df %>%
  dplyr::filter(is_living_with_dog == FALSE) %>%
  dplyr::count(curr_dog_location) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

## Expectations and Considerations

Generate breakdown for meeting of expectations.

```{r}
df %>%
  dplyr::count(met_expectations) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

```{r}
df %>%
  dplyr::group_by(is_satisfied) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Get count of owners with at least one dog that met expectations.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  dplyr::summarise(sat_cnt = sum(is_satisfied)) %>%
  dplyr::filter(sat_cnt > 0)
```

Get a count of owners where all dogs met expectations.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  dplyr::summarise(sat_cnt = sum(is_satisfied), tot_cnt=n()) %>%
  dplyr::filter(sat_cnt == tot_cnt)
```

Generate breakdown of acquisition sources.

```{r}
df %>%
  dplyr::count(acquisition_source) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Generate breakdown for time spent thinking.

```{r}
df %>%
  dplyr::count(time_thinking) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Generate breakdown for primary motivation.

```{r}
df %>%
  dplyr::count(primary_motivation) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Generate a summary for characteristic ranks.

```{r}
df %>%
  dplyr::select(contains("rank") & !contains("revised")) %>%
  summary()
```

## Consideration of Another Dog

Breakdown of owners that would consider getting another dog.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  tidyr::fill(is_consider_another_dog, .direction="downup") %>%
  dplyr::ungroup() %>%
  dplyr::distinct(owner_id, .keep_all=TRUE) %>%
  dplyr::count(is_consider_another_dog) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Breakdown of fully satisfied owners considering another dog.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  dplyr::mutate(sat_cnt = sum(is_satisfied), tot_cnt=n()) %>%
  dplyr::filter(sat_cnt == tot_cnt) %>%
  tidyr::fill(is_consider_another_dog, .direction="downup") %>%
  dplyr::ungroup() %>%
  dplyr::distinct(owner_id, .keep_all=TRUE) %>%
  dplyr::count(is_consider_another_dog) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Breakdown of owners satisfied with some, but not all, dogs considering another
dog.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  dplyr::mutate(sat_cnt = sum(is_satisfied), tot_cnt=n()) %>%
  dplyr::filter(sat_cnt < tot_cnt, sat_cnt > 0) %>%
  tidyr::fill(is_consider_another_dog, .direction="downup") %>%
  dplyr::ungroup() %>%
  dplyr::distinct(owner_id, .keep_all=TRUE) %>%
  dplyr::count(is_consider_another_dog) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Breakdown of owners not satisfied with any dogs considering another dog.

```{r}
df %>%
  dplyr::group_by(owner_id) %>%
  dplyr::mutate(sat_cnt = sum(is_satisfied), tot_cnt=n()) %>%
  dplyr::filter(sat_cnt == 0) %>%
  tidyr::fill(is_consider_another_dog, .direction="downup") %>%
  dplyr::ungroup() %>%
  dplyr::distinct(owner_id, .keep_all=TRUE) %>%
  dplyr::count(is_consider_another_dog) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Breakdown of revised acquisition source.

```{r}
df %>%
  dplyr::group_by(revised_acquisition_source) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = round(n / sum(n) * 100, 2)) %>%
  janitor::adorn_totals("row")
```

Generate a summary for revised characteristic ranks.

```{r}
df %>%
  dplyr::select(contains("rank") & contains("revised")) %>%
  summary()
```

Difference in rank between revised rank and average of prior ranks.

```{r}
df %>%
  dplyr::select(contains("rank")) %>%
  dplyr::summarise(
    dplyr::across(
      everything(),~ janitor::round_half_up(mean(., na.rm = TRUE), 2))) %>%
  dplyr::mutate(
    age_delta = revised_age_rank - age_rank,
    app_delta = revised_appearance_rank - appearance_rank,
    breed_delta = revised_breed_rank - breed_rank,
    comp_delta = revised_compatability_rank - compatability_rank,
    pers_delta = revised_personality_rank - personality_rank,
    size_delta = revised_size_rank - size_rank,
    train_delta = revised_trainability_rank - trainability_rank) %>%
  summary()
```

# Data Exploration

## Numeric Correlations

```{r, fig.width=5, fig.height=4}
pairs.panels(df[,12:19])
```

## Fields Split by Met Expecations

### Owner Fields

Split by owner gender.

```{r}
df %>%
  dplyr::count(owner_gender, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::group_by(owner_gender) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  ggplot(aes(x=reorder(owner_gender, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("owner gender") +
    ylab("frequency")
```

Split by owner age.

```{r}
means <- df %>%
  dplyr::group_by(is_satisfied) %>%
  dplyr::summarise(means = mean(owner_age))
df %>%
  ggplot(aes(owner_age, fill = is_satisfied)) +
    geom_histogram(alpha=0.7, position="identity", aes(y = ..density..),
                   color="black", bins=30) +
    geom_density(alpha=0.7) +
    geom_vline(xintercept=means$means, color=c("#F8766D", "#00BFC4"),
               linetype="dashed", size=0.75)
```

### Dog Fields

Split by acquisition source.

```{r}
df %>%
  dplyr::count(acquisition_source, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::group_by(acquisition_source) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  dplyr::mutate(
    acquisition_source = fct_recode(
      acquisition_source,
      "family/friend"="family_or_friend",
      "pet shop"="pet_shop",
      "shelter/rescue"="rescue"
    )
  ) %>%
  ggplot(aes(x=reorder(acquisition_source, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("acquisition source") +
    ylab("frequency") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Split by living situation.

```{r}
df %>%
  dplyr::count(is_living_with_dog, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::group_by(is_living_with_dog) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  ggplot(aes(x=reorder(is_living_with_dog, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("living with dog") +
    ylab("frequency")
```

## Considerations Fields

Split by time spent thinking.

```{r}
df %>%
  dplyr::count(time_thinking, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::group_by(time_thinking) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  ggplot(aes(x=reorder(time_thinking, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("time spent thinking") +
    ylab("frequency")
```

Split by primary motivation.

```{r}
df %>%
  dplyr::count(primary_motivation, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::group_by(primary_motivation) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  ggplot(aes(x=reorder(primary_motivation, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("primary motivation") +
    ylab("frequency") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Split by consideration of another dog.

```{r}
df %>%
  dplyr::filter(!is.na(is_consider_another_dog)) %>%
  dplyr::count(is_consider_another_dog, is_satisfied) %>%
  tidyr::spread(is_satisfied, n) %>%
  janitor::adorn_totals("row") %>%
  janitor::adorn_totals("col")
```

```{r}
df %>%
  dplyr::filter(!is.na(is_consider_another_dog)) %>%
  dplyr::group_by(is_consider_another_dog) %>%
  dplyr::mutate(freq = sum(is_satisfied) / n()) %>%
  ggplot(aes(x=reorder(is_consider_another_dog, desc(freq)), fill=is_satisfied)) +
    geom_bar(position="fill") +
    xlab("would consider an additional dog") +
    ylab("frequency") +
    theme(axis.text.x = element_text(angle = 45, hjust=1))
```

Split by ranks.

```{r, fig.width=5, fig.height=3}
featurePlot(
  x = df[, 13:19],
  y = factor(df$is_satisfied),
  plot = "box",
  scales = list(y = list(relation="free"), x = list(rot = 90)),
  layout = c(4,2),
  auto.key = list(columns = 2)
)
```

Split by revised ranks.

```{r, fig.width=5, fig.height=3}
featurePlot(
  x = df[, 20:26],
  y = factor(df$is_satisfied),
  plot = "box",
  scales = list(y = list(relation="free"), x = list(rot = 90)),
  layout = c(4,2),
  auto.key = list(columns = 2)
)
```

Split by source.

```{r, fig.width=5, fig.height=3}
featurePlot(
  x = df[, 13:19],
  y = factor(df$acquisition_source),
  plot = "box",
  scales = list(y = list(relation="free"), x = list(rot = 90)),
  layout = c(4,2),
  auto.key = list(columns = 2)
)
```

## Contingency Tables

```{r}
table(df$primary_motivation, df$time_thinking)
```

```{r}
table(df$acquisition_source, df$time_thinking)
```

## Mosiac Plots

Plot of background variables to be used in model.

```{r}
#mosaic(~ owner_gender + primary_motivation + is_satisfied, data=df,
#       highlighting = "is_satisfied", highlighting_fill = c("lightblue", "pink"),
#       direction = c("h", "v", "v"))
```

## Violin Plots

```{r}
df_long <- df %>%
  dplyr::select(owner_gender, is_satisfied, dplyr::contains("rank"), time_thinking) %>%
  rename_with(~ gsub("revised_(.*)", "\\1_revised", .)) %>%
  tidyr::pivot_longer(
    cols=age_rank:trainability_rank_revised,
    names_to=c("char", "revised"),
    names_pattern="(.*)_rank_?(revised)?",
    values_to="rank",
    values_drop_na=TRUE
  ) %>%
  dplyr::mutate(
    Adoption=factor(ifelse(revised =="revised", "future", "past"),
                    levels=c("past", "future"))) %>%
  dplyr::mutate(char=as.factor(char))

width <- 5
df_long %>%
  ggplot(aes(x=char, y=rank, fill=Adoption)) +
    introdataviz::geom_split_violin(alpha=.5, trim=T) +
    geom_boxplot(width=.125, alpha=.8, show.legend=FALSE) +
    scale_x_discrete(name = "Characteristic") +
    scale_y_continuous(name = "Rank",
                       limits = c(1, 7),
                       breaks = seq(1, 7, 1))
ggsave("fig-1-violin-plot.pdf")
```

## Ridge Plots

```{r}
df_long %>%
  dplyr::filter(Adoption=="past") %>%
  ggplot(aes(x=rank, y=char, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
    stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
    scale_fill_viridis_c(name = "Tail probability", direction = -1)
```
```{r}
df_long %>%
  dplyr::filter(Adoption=="future") %>%
  ggplot(aes(x=rank, y=char, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
    stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
    scale_fill_viridis_c(name = "Tail probability", direction = -1)
```

```{r}
df_long %>%
  ggplot(aes(x=rank, y=char, fill=Adoption,
             group=interaction(char, !ifelse(Adoption=="past", F, T)))) +
    ggridges::geom_density_ridges(alpha=0.5) +
  theme_ridges() +
  labs(
    x = "Rank",
    y = "Characteristic",
  )
ggsave("fig-1-ridgeline-plot.pdf")
```
## Jitter Plots

```{r}
df_long2 <- df_long %>%
  mutate(char2 = 2*as.numeric(as.factor(char))+0.8*as.numeric(as.factor(Adoption)))

df_long2 %>%
  ggplot(aes(x=char2, y=rank)) +
    geom_jitter(aes(color=Adoption), width=0.3, alpha=0.3) +
    geom_boxplot(aes(x=char2, y=rank,
                     group=interaction(char, Adoption)),
                 width=0.35, alpha=0.1) +
    scale_x_continuous(breaks=unique(df_long2$char2[1:7])+.4,
                       labels=unique(df_long2$char)) +
    labs(
      y = "Rank",
      x = "Characteristic",
    ) +
    guides(colour = guide_legend(override.aes = list(alpha = 1)))
ggsave("fig-1-jitter-plot.pdf")
```

## Box Plots

```{r}
df_long2 <- df_long %>%
  mutate(char2 = 2*as.numeric(as.factor(char))+0.8*as.numeric(as.factor(Adoption)))

df_long2 %>%
  ggplot(aes(x=char2, y=rank)) +
    #geom_jitter(aes(color=Adoption), width=0.3, alpha=0.3) +
    geom_boxplot(aes(x=char2, y=rank,
                     group=interaction(char, Adoption),
                     color=Adoption),
                 width=0.35, alpha=0.1) +
    scale_x_continuous(breaks=unique(df_long2$char2[1:7])+.4,
                       labels=unique(df_long2$char)) +
    labs(
      y = "Rank",
      x = "Characteristic",
    ) +
    guides(colour = guide_legend(override.aes = list(alpha = 1)))
ggsave("fig-1-box-plot.pdf")
```

```{r}
df_long2 %>%
  ggplot(aes(x=char, y=rank, fill=Adoption)) +
  geom_boxplot() +
  labs(
    y = "Rank",
    x = "Characteristic",
  )
ggsave("fig-1-box-plot.pdf")
```

## Nested Bar Chart

```{r}
df_long_totals <- df_long %>%
  dplyr::select(owner_gender, is_satisfied, time_thinking) %>%
  plyr::ddply(.(owner_gender, time_thinking), summarize, is_satisfied=sum(is_satisfied))


titanic.colors<-list("gray90",c("#0000ff","#7700ee","#aa00cc","#dd00aa"),
c("#ddcc00","#ee9900"),c("pink","lightblue"))

barNest(
  is_satisfied~owner_gender+time_thinking,
  data=df_long_totals,
  col=titanic.colors,
  #showall=TRUE,
  main="Owner satisfaction by owner gender, age, and forethought",
  ylab="Proportion satisfied"
  #FUN=c("propbrk")#,"binciWu","binciWl","valid.n"),
  #shrink=0.15
  #trueval="Yes"
)
```

## Principal Component Analysis

```{r}
df_pca <- df %>%
  dplyr::select(contains("rank") & !contains("revised"))
pc <- prcomp(df_pca, center = TRUE, scale = TRUE)
attributes(pc)
print(pc)
stats::biplot(pc, scale = 0)
# Explained variance.
round(pc$sdev^2 / sum(pc$sdev^2), 2)
```

```{r}
g <- ggbiplot(pc,
              obs.scale = 1,
              var.scale = 1,
              groups = df$is_satisfied,
              ellipse = TRUE,
              circle = TRUE,
              ellipse.prob = 0.68)
g <- g + scale_color_discrete(name = '')
g <- g + theme(legend.direction = 'horizontal',
               legend.position = 'top')
print(g)
```

```{r}
pc <- PCAtools::pca(df_pca)
PCAtools::screeplot(pc)
```
```{r, fig.width=3, fig.height=3}
PCAtools::biplot(pc, showLoadings = T)
```

```{r, fig.height=4}
PCAtools::plotloadings(pc, labSize = 3)
```

```{r, fig.width=5, fig.asp=1}
pairsplot(
  pc,
  triangle = TRUE,
  trianglelabSize = 12,
  hline = 0,
  vline = 0,
  pointSize = 1
)
```

## Factor Analysis

```{r}
df_rank <- df[,c(13, 18:19)]
fa(r=cor(df_rank), nfactors=dim(df_rank)[2], rotate="varimax", SMC=FALSE, fm="pa")
```

# Save Session Info

```{r}
sessionInfo()
```

# Cleanup

```{r}
rm(list = ls())
```
